---
title: "homework-3"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{homework-3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup}
library(bis557)
```


QUESTION 1
```{r}
set.seed(1234)
X <- matrix(rnorm(9), nrow = 3, ncol = 3)
XX <- t(X) %*% X
svd(XX)$d

p <- c(0.9999, 0.1, 1e-15)
D <- p * (1-p)
H <- t(X) %*% diag(D) %*% X
svd(H)$d
```
From the above results we can know the linear Hessian $X^TX$ is well conditioned, as the ratio of the largest singular value to the smallest value is approximately 61. However the longistic variation has the ratio around $10^{14}$, which means it is ill-conditioned.



QUESTION 2

glm_new() implements a first-order solution for the GLM maximum likelihood problem using only gradient information. In the following examples, a constant step size and a Momentum adaptive step size are used to fit the GLM model.
```{r}
library(bis557)

#devtools::build_vignettes()
n <- 5000
p <- 3
beta <- c(-1, 0.2, 0.1)
X <- cbind(1, matrix(rnorm(n * (p- 1)), ncol = p - 1))
eta <- X %*% beta
lambda <- exp(eta)
y <- rpois(n, lambda = lambda)
#Example 1
fit_linear_model <- glm_new(X, y, mu_fun = function(eta) exp(eta), var_fun = identity,
                            step = 0.0001, method = "constant")
fit_linear_model$coefficients
#Example 2
fit_linear_model2 <- glm_new(X, y, mu_fun = function(eta) exp(eta), var_fun = identity, step = 0.0001, 
                             friction = 0.9, method = "momentum")
fit_linear_model2$coefficients
```


QUESTION 3

The function multi_class() is built to classify multi-categorical variables based on logistic regression. It returns the class with the highest probability and the corresponding coefficients.
```{r}
library(bis557)
data(iris)
X <- cbind(1, as.matrix(iris[, -5]))
y <- ifelse(iris$Species==levels(iris$Species)[1], 1, ifelse(iris$Species==levels(iris$Species)[2], 2, 3))

fit_linear_model <- multi_class(X, y, itr = 50, tolerance = 1e-10)
fit_linear_model$class
fit_linear_model$coefficients
```


REFERENCE

Textbook CASL P124 - 138.








